---
title: "Final Project Rmarkdown"
author: "Kevin Stutenberg"
date: "December 5, 2018"
output: html_document
---


We can connect to MySQL using the RMySQL package.
# ```{r}
# install.packages("RMySQL");
# install.packages("dbplyr");
# install.packages("dplyr");
# install.packages("ggplot2");
# install.packages("knitr");
# ```

Initialize libraries
```{r, echo=FALSE}
library(RMySQL);
library(dplyr);
library(ggplot2);
library(DBI);
library(tidyverse)
```

Load the data using the readr package, or alternatively enter the path to the data, and load it within r
# ```{r}
# datapath= "D:/Taxi_Trips_2016.csv"
# 
# Taxi_Trips <- read_csv("D:/Taxi_Trips_2016.csv", 
# +     col_types = cols(`Dropoff Census Tract` = col_skip(), 
# +         `Dropoff Centroid  Location` = col_skip(), 
# +         `Pickup Census Tract` = col_skip(), 
# +         `Pickup Centroid Location` = col_skip(), 
# +         `Taxi ID` = col_skip(), `Trip ID` = col_skip()))
# ```

Filter the data to remove rows which have no values in Time stamp or location values
```{r}
Taxi_Trips_reduced<- na.omit(Taxi_Trips, cols=c("Trip Start Timestamp", "Trip End Timestamp", "Pickup Centroid Latitude",  "Pickup Centroid Longitude",  "Dropoff Centroid Latitude",  "Dropoff Centroid Longitude"));
                             
head(Taxi_Trips_reduced)
```

Create a new column with the id for verification
```{r}
Taxi_Trips_reduced$New_ID <- seq.int(nrow(Taxi_Trips_reduced))
head(Taxi_Trips_reduced, 50)
```


Test Upload - When needed to test a small upload
```{r}
test<-head(Taxi_Trips_reduced,200)
test
```

Set up the connection to AWS
```{r, echo=FALSE}
connection = dbConnect(MySQL(),user="msca", password="yessqlnosql", 
                dbname="divvybikes", host="mscadbinstance.cxqhidondkea.us-east-1.rds.amazonaws.com");
```

Upload the new table 
```{r}
dbWriteTable(connection, name='Taxi_2015', value=Taxi_Trips_reduced)
```

```{r}
myQuery <- "SELECT COUNT(*) FROM Taxi_2016_test;"
dbGetQuery(connection, myQuery)
```
  

Close active DB Connections
```{r}
dbGetQuery(connection, "show processlist")

dbDisconnectAll <- function()
{
  ile <- length(dbListConnections(MySQL())  )
  lapply( dbListConnections(MySQL()), function(x) dbDisconnect(x) )
  cat(sprintf("%s connection(s) closed.\n", ile))
}

dbDisconnectAll()

```






